Distributed Range-Only Localisation that Preserves Sensor and Navigator Privacies

### 2021-06-04 main.tex
Highlighted novelty of the work more than before (as per reviewer request) Explicitly pointed out how the accuracy of the filter can be assessed form the provided figures (accuracy was said to be missing by a reviewer) Fixed problem with the reasoning behind taking a 95% confident conservative modified measurement variance estimate (error pointed out by reviewer)

Our contribution in this work considers range-only localisation with formal cryptographic requirements ensuring that sensors keep their measurements and locations private and that navigator estimates cannot be learned by sensors. We define a novel notion for private linear combination aggregation, satisfying these cryptographic requirements, before using it to derive a filter based on the Extended Kalman Filter, with no hierarchical sensor layout assumptions. The linear combination aggregation scheme put forward is in principle similar to the pWSAh scheme in \cite{alexandruPrivateWeightedSum2020}, however, a formal definition capturing communication assumptions and leakages is given, crucial for cryptographic security.

We motivate this scenario with the example of vehicle localisation in the presence of privately owned measurement stations. While the intention of measurement stations is the accurate navigation of passing vehicles, it may be reasonable to desire identifying location or hardware details to remain unknown at other present stations and to the navigator. Similarly, a navigator may not wish to disclose their most accurate location estimates to untrusted third-party measurers.

@@

Computing $z'_{k,i}$ and its variance $r'_{k,i}$ from the original measurements $z_{k,i}$ are complicated by the noise term $v_{k,i} \sim \mathcal{N}(0, r_{k,i})$, and simply squaring the original range measurements produces
\begin{equation}
    \begin{split}
        z_{k,i}^2 &= (h_i(\vec{x}_k) + v_{k,i})^2 \\
        &= h'_i(\vec{x}_k) + 2h_i(\vec{x}_k)v_{k,i} + v_{k,i}^2\,,
    \end{split}
\end{equation}
with a new noise term $2h_i(\vec{x}_k)v_{k,i} + v_{k,i}^2$, now dependent on the measurement function $h_i$, and no longer zero-mean. We can compute the mean of this new noise term (a function of the Gaussian term $v_{k,i}$) as
\begin{equation}
    \mathsf{E}[2h_i(\vec{x}_k)v_{k,i} + v_{k,i}^2] = r_{k,i}\, \label{eqn:modified_noise_mean}
\end{equation}
and mean-adjust modified measurements as
\begin{equation}
    \begin{split}
        z'_{k,i} &= z_{k,i}^2 - r_{k,i} \\
        &= h_i(\vec{x}_k)^2 + 2h_i(\vec{x}_k)v_{k,i} + v_{k,i}^2 - r_{k,i} \\
        &= h'_i(\vec{x}_k) + v'_{k,i}\,,
    \end{split} \label{eqn:modified_measurement}
\end{equation}
with now zero-mean noise $v'_{k,i} = 2h_i(\vec{x}_k)v_{k,i} + v_{k,i}^2 - r_{k,i}$. The noise in this case (again a function of Gaussian term $v_{k,i}$) has variance 
\begin{equation}
    \mathsf{Var}[v'_{k,i}] = 4h_i(\vec{x}_k)^2r_{k,i} + 2r_{k,i}^2
\end{equation}
and is also dependent on $h_i$. To use the modified measurement \eqref{eqn:modified_measurement} with the EIF, we require an estimate for $\mathsf{Var}[v'_{k,i}]$ at the sensor as well. Additionally, a conservative estimate (\textit{i.e.}, a larger variance resulting in less confidence in measurements) is desirable to reduce filter divergence. The naive approach, replacing $h_i(\vec{x}_k)$ with $z_{k,i}$, only provides a conservative estimate when
\begin{equation}
    \begin{split}
        &z_{k,i} > h_i(\vec{x}_k) \\
        \implies &v_{k,i} > 0\,,
    \end{split}
\end{equation}
which cannot be guaranteed. Instead, we provide a conservative estimate with $95\%$ confidence due to the Gaussianity of $v_{k,i}$, by adjusting the replacement term $z_{k,i}$ by two of its standard deviations $\sqrt{r_{k,i}}$. The modified measurement's variance at timestep $k$ is therefore conservatively approximated by
\begin{equation}
    \begin{split}
        r'_{k, i} &= 4(z_{k,i} + 2\sqrt{r_{k,i}})^2r_{k,i} + 2r_{k,i}^2 \\
        &\gtrapprox \mathsf{Var}[v'_{k,i}]
    \end{split}\,. \label{eqn:modified_measurement_variance_estimate}
\end{equation}
at each sensor $i$.

@@

We have considered multiple sensor layouts, each with four sensors, to capture the dependence of estimated modified measurement variances $r'_{k,i}$ on the original measurements $z_{k,i}$. The layouts, of varying sensor distances to the navigator track, are shown in figure \ref{fig:sim_layouts}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layouts.pdf}
    \caption{Diffrent simulation layouts with varying average distances between navigator and sensors.}
    \label{fig:sim_layouts}
\end{figure}
The simulation root mean square error (RMSE) of the privacy-preserving method has been compared to the RMSE of the standard EIF (an algebraic equivalent of the EKF), using unmodified measurements. Measurement and estimation in each layout from figure \ref{fig:sim_layouts} consists of $50$ filter iterations and was run $100$ times. Unmodified measurement variances were taken as $r_{k,i}=5$ for all $k>0$ and a large fractional precision factor, $\phi=2^{32}$, was chosen. The results can be seen in figure \ref{fig:sim_layout_errors}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layout_errors.pdf}
    \caption{MSE of the privacy-preserving EIF and standard EIF for different layout simulations.}
    \label{fig:sim_layout_errors}
\end{figure}
From these results, we can see the similarity in filter performance between the privacy-preserving method and that of the traditional EIF. We can also see that the varying average distances between sensors and navigator have little impact on the differences in performance. We can attribute the similarity in MSE to the conservativeness of estimated modified measurement variances $r'_{k,i}$, resulting in few additional filter divergences, and to the high fractional precision factor, keeping computations consistent with the floating-point arithmetic of the EIF.

>>>>

Our contribution in this work presents a range-only localisation method meeting formal cryptographic requirements that ensure sensors keep their measurements and locations private and navigator estimates cannot be learned by sensors. We first define a novel notion for private linear combination aggregation, satisfying these cryptographic requirements, before using it to derive a filter based on the Extended Kalman Filter, with no hierarchical sensor layout assumptions. The linear combination aggregation scheme we put forward is in principle similar to the pWSAh scheme in \cite{alexandruPrivateWeightedSum2020}, however, a formal definition capturing communication assumptions and leakages is given, crucial for cryptographic security. To the best of the authors' knowledge, no existing method for Bayesian state estimation using range-only sensors and meeting the desired privacy requirements exists.

We motivate this scenario with the example of vehicle localisation in the presence of privately owned measurement stations. While the intention of measurement stations is the accurate navigation of passing vehicles, it may be reasonable to desire identifying location or hardware details to remain unknown to the other present stations and to the navigator. Similarly, a navigator may not wish to disclose their most accurate location estimates to untrusted third-party measurers.

@@

Computing $z'_{k,i}$ and its variance $r'_{k,i}$ from the original measurements $z_{k,i}$ are complicated by the noise term $v_{k,i} \sim \mathcal{N}(0, r_{k,i})$, and simply squaring the original range measurements produces
\begin{equation}
    \begin{split}
        z_{k,i}^2 &= (h_i(\vec{x}_k) + v_{k,i})^2 \\
        &= h'_i(\vec{x}_k) + 2h_i(\vec{x}_k)v_{k,i} + v_{k,i}^2\,,
    \end{split}
\end{equation}
with a new noise term $2h_i(\vec{x}_k)v_{k,i} + v_{k,i}^2$, now dependent on the measurement function $h_i$, and no longer zero-mean. We can compute the mean of this new noise term (a function of the Gaussian term $v_{k,i}$) as
\begin{equation}
    \mathsf{E}[2h_i(\vec{x}_k)v_{k,i} + v_{k,i}^2] = r_{k,i}\, \label{eqn:modified_noise_mean}
\end{equation}
and mean-adjust modified measurements as
\begin{equation}
    \begin{split}
        z'_{k,i} &= z_{k,i}^2 - r_{k,i} \\
        &= h_i(\vec{x}_k)^2 + 2h_i(\vec{x}_k)v_{k,i} + v_{k,i}^2 - r_{k,i} \\
        &= h'_i(\vec{x}_k) + v'_{k,i}\,,
    \end{split} \label{eqn:modified_measurement}
\end{equation}
with now zero-mean noise $v'_{k,i} = 2h_i(\vec{x}_k)v_{k,i} + v_{k,i}^2 - r_{k,i}$. The noise in this case (again a function of Gaussian term $v_{k,i}$) has variance 
\begin{equation}
    \mathsf{Var}[v'_{k,i}] = 4h_i(\vec{x}_k)^2r_{k,i} + 2r_{k,i}^2 \label{eqn:modified_measurement_variance}
\end{equation}
and is also dependent on $h_i$. To use the modified measurement \eqref{eqn:modified_measurement} with the EIF, we require an estimate for $\mathsf{Var}[v'_{k,i}]$ at the sensor as well. Additionally, a conservative estimate (\textit{i.e.}, a larger variance resulting in less confidence in measurements) is desirable to reduce filter divergence. While the naive approach, replacing $h_i(\vec{x}_k)$ with $z_{k,i}$ in \eqref{eqn:modified_measurement_variance}, may not provide a conservative estimate when $z_{k,i}^2 < h_i(\vec{x}_k)^2$, the Gaussianity of $v_{k,i}$ can be exploited to provide a conservative estimate with $95\%$ confidence by shifting the replacement term $z_{k,i}$ by two of its standard deviations $\sqrt{r_{k,i}}$. The modified measurement's variance at timestep $k$ can therefore be conservatively approximated by
\begin{equation}
    \begin{split}
        r'_{k, i} &= 4(z_{k,i} + 2\sqrt{r_{k,i}})^2r_{k,i} + 2r_{k,i}^2 \\
        &\gtrapprox \mathsf{Var}[v'_{k,i}]\,,
    \end{split} \label{eqn:modified_measurement_variance_estimate}
\end{equation}
at each sensor $i$.

@@

We have considered multiple sensor layouts, each with four sensors, to capture the dependence of estimated modified measurement variances $r'_{k,i}$ on the original measurements $z_{k,i}$. The layouts, of varying sensor distances to the navigator track, are shown in figure \ref{fig:sim_layouts}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layouts.pdf}
    \caption{Diffrent simulation layouts with varying average distances between navigator and sensors.}
    \label{fig:sim_layouts}
\end{figure}
To demonstrate the accuracy of the method, we have compared the root mean square error (RMSE) of the privacy-preserving filter to the standard EIF using unmodified measurements, that is algebraically equivalent to the EKF, typically used in industry for linearising non-linear state estimation. Measurement and estimation in each layout from figure \ref{fig:sim_layouts} consists of $50$ filter iterations and was run $100$ times. Unmodified measurement variances were taken as $r_{k,i}=5$ for all $k>0$ and a large fractional precision factor, $\phi=2^{32}$, was chosen. The results can be seen in figure \ref{fig:sim_layout_errors}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layout_errors.pdf}
    \caption{MSE of the privacy-preserving EIF and standard EIF for different layout simulations.}
    \label{fig:sim_layout_errors}
\end{figure}
From these results, we can see the similarity in filter performance between the privacy-preserving method and that of the traditional EIF. We can also see that the varying average distances between sensors and navigator have little impact on the differences in performance. We can attribute the similarity in MSE to the conservativeness of estimated modified measurement variances $r'_{k,i}$, resulting in few additional filter divergences, and to the high fractional precision factor, keeping computations consistent with the floating-point arithmetic of the EIF.
